Water desalination using polyelectrolyte hydrogel.

### 2022-09-03 Creta_ecis2022/presentation/beamer.tex
presentation finalized (almost)

\section[Intro]{Introduction}
\begin{frame}[fragile]{Introduction. Desalination.}
\begin{enumerate}
\item Distillation
\item Reverse osmosis
\item Forward osmosis
\end{enumerate}
\end{frame}
\begin{frame}[fragile]{Introduction. Forward osmsis.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]{Introduction. Manfred Wilhelm experiment}
	\begin{columns}[T,onlytextwidth]% COLUMNS
		\column{0.4\textwidth}
			\begin{figure}[t]
				{\includegraphics[width=0.9\columnwidth]{{figures/exp}.png}}
			\end{figure}

		\column{0.6\textwidth}
			\begin{figure}[t]
				{\includegraphics[width=0.9\columnwidth]{{figures/scheme}.png}}
			\end{figure}
	\end{columns}% COLUMNS

@@

@@

\subsection{Grand-reaction ensemble.}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{figures/grand_reaction_model.png}}
\begin{frame}{Grand-reaction ensemble.}
\begin{columns}[T,onlytextwidth]% COLUMNS
\column{0.3\textwidth}
\vspace{4.5cm}
Ionization equilibrium in \textbf{reaction} ensemble
\begin{equation*}
	\boxed{
		\mathrm{pA^- + H_3O^+ \leftrightarrows pAH}
	}
\end{equation*}

\column{0.4\textwidth}
Constant salinity 
\begin{equation*}
\boxed{c_s = c_\mathrm{Cl}^- = c_\mathrm{Na}^+}
\end{equation*}
The simulation box freely 
exchanges ions with reservoir as in \textbf{grand cannonical} ensemble
\end{columns} % COLUMNS
\vspace{0.5cm}
{\scriptsize Landsgesell, J., Hebbeker, P., Rud, O., Lunkad, R., Košovan, P., \\ Holm, C. (2020). 
{\bf Grand-Reaction Method for Simulations of Ionization \\
Equilibria Coupled to Ion Partitioning. }
Macromolecules, 53(8), 3007–3020. }
\end{frame}
}

{

}

{

\begin{frame}{Grand-reaction ensemble.}

	\begin{columns}[T,onlytextwidth]% COLUMNS
		\column{0.5\textwidth}
			\begin{figure}
				\includegraphics[height=4.cm]{figures/standard_states.pdf}
			\end{figure} 
		\column{0.5\textwidth}	
			\textbf{Reaction ensemble}\\ 
			\vspace{1cm}
			The reaction of an acidic unit 
			\begin{equation*}
				\mathrm{HA}\stackrel{K}{\leftrightarrows}\mathrm{A^-}+\mathrm{H^+}\label{eq: acid reaction}
			\end{equation*}
			\begin{equation*}
				\Omega=E-TS+\sum_i\left(\mu_i-\mu^\ominus_i\right) N_i\label{eq:Omega-RE}
			\end{equation*}
	\end{columns} % COLUMNS
	\vspace{1cm}
	Then the change of system free energy during a single reaction step
	\begin{equation*}
		\Delta\Omega=k_BT\ln\left(\prod_i V^{\nu_i\xi}\frac{N_i!}{\left(N_i+\nu_i\xi\right)!}\right)+
		\xi\left(\sum_i\nu_i\mu_i-\sum_i\nu_i\mu_i^\ominus\right)+\Delta E
		\label{eq: DeltaG RE one}
	\end{equation*}
	{\tt accept if $\mathcal{R}^{\xi}<e^{\Delta\Omega/k_BT}$}
\end{frame}
}

\begin{frame}{Grand-reaction ensemble.}
\begin{columns}[T,onlytextwidth]% COLUMNS
\column{0.7\textwidth}
\begin{equation*}\boxed{
\Delta\Omega=k_BT\ln\left(K^\xi\prod_i V^{\nu_i\xi}\frac{N_i!}{\left(N_i+\nu_i\xi\right)!}\right)+\Delta E}
\end{equation*}
\column{0.3\textwidth}
\begin{equation*}\boxed{
K = e^{-\sum_i\nu_i \mu_i^\ominus }} 
\end{equation*}
\end{columns} % COLUMNS

\vspace{0.6cm}
\begin{columns}[T,onlytextwidth]% COLUMNS
\column{0.5\textwidth}
	\begin{equation*}
		\mathrm{HA}\stackrel{K}{\leftrightarrows}\mathrm{A^-}+\mathrm{H^+} 
	\end{equation*}
	\begin{equation*}
		K= \mu_\h^\ominus+\mu_\mathrm{A^-}^\ominus - \mu_\mathrm{HA}^\ominus
	\end{equation*}
	\vspace{0.2cm}
\begin{equation*}
\emptyset \leftrightarrows  \na  +\cl
\end{equation*}
\begin{equation*}
K= \mu_\na+\mu_\cl
\end{equation*}
\vspace{0.2cm}
\begin{equation*}
\emptyset \leftrightarrows \ca + 2\cl
\end{equation*}
\begin{equation*}
K= 2\mu_\ca+\mu_\cl
\end{equation*}
\column{0.5\textwidth}
			\begin{figure}
				\includegraphics[height=4.cm]{figures/standard_states.pdf}
			\end{figure} 
\end{columns} % COLUMNS

\begin{frame}{Simulation protocol.}
\begin{enumerate}
\item {\tt Choose randomly: LMD, RE, or EX.}	\vspace{0.5cm}
\item {\tt Simulate the chosen, collecting 50 samples of: }
	\vspace{0.3cm}
	\begin{columns}[T,onlytextwidth]% COLUMNS
		\column{0.3\textwidth}
			\fbox{\parbox{\textwidth}{LMD: pressure, $P$, and $\{R_e\}$}}
		\column{0.3\textwidth}
	       		\fbox{\parbox{\textwidth}{RE: number of ionized segments, $\NA$}}
      		\column{0.3\textwidth}
        		\fbox{\parbox{\textwidth}{EX: number of salt ions, $N_\na$ and $N_\cl$}}
	\end{columns} % COLUMNS
	\vspace{0.5cm}
    {\tt Check the autocorrelation of each samples array.\\
    Pearson coefficient must be < 0.2. }
	\vspace{0.5cm}
\item {\tt Repeat collecting at least 200 averages from each process}.
\end{enumerate}
\end{frame}

@@

\begin{frame}{Monovalent salt. Compression.}
	\begin{figure}[t]
		\subfloat[low salinity, $c_s = 0.007$ mol/l]
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_pressure_log_density_cs0.006_N_Samples200}.pdf}}
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_alpha_log_density_cs0.006_N_Samples200}.pdf}}
		\subfloat[high salinity, $c_s = 0.209$ mol/l]
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_pressure_log_density_cs0.15_N_Samples200}.pdf}}
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_alpha_log_density_cs0.15_N_Samples200}.pdf}}
	\end{figure}
\end{frame}

\begin{frame}{Monovalent salt. ``No electrostatics'' vs ``Mean field theory''.}
	\begin{equation*}
		\frac{\alpha}{1-\alpha}10^{\pK-\pH} =\sqrt{1+\left(\frac{\alpha\cp}{2\cs}\right)^{2}} - \frac{\alpha\cp}{2\cs}
	\end{equation*}
	Together with electroneutrality condition it translates to
	\begin{equation*}
		-\frac{{{\alpha}^{3}}{\cp}}{{\cs}}+
		{\alpha}^{2}\left(\frac{{\cp}}{{\cs}}+\Theta-\frac{1}{\Theta}\right)+
		\frac{2 \alpha}{\Theta}-\frac{1}{\Theta}=0
	\end{equation*}
	where $\Theta = 10^{\pK-\pH}$.
	\begin{figure}
		\subfloat[low salinity, $c_s = 0.007$ mol/l]
			{\includegraphics[width=0.47\columnwidth]{{figures/gel_alpha_log_density_lB0.0_cs0.006_N_Samples200}.pdf}}
\subfloat[high salinity, $c_s = 0.209$ mol/l]
			{\includegraphics[width=0.47\columnwidth]{{figures/gel_alpha_log_density_lB0.0_cs0.15_N_Samples200}.pdf}}
	\end{figure}
\end{frame}

\begin{frame}{Monovalent salt. Desalination effect.}

	\begin{figure}[t]
		{\includegraphics[width=0.5\columnwidth]{{figures/gibbs3}.pdf}}
		\caption{
			Schematic illustration of the hydrogel being compressed isotropically by pressure $\Delta P$ 
			from the initial volume at free swelling equilibrium, $V_0$ to a volume $\Vgel$. 
			Simultaneously, the gel exchanges small ions with a reservoir solution of salinity $\cs$.}
	\end{figure}
\end{frame}

\begin{frame}{Monovalent salt. Desalination effect.}
	\begin{figure}[t]
		\subfloat[low salinity, $c_s = 0.007$ mol/l]
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_Na_log_density_cs0.006_N_Samples200}.pdf}}
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_Cl_log_density_cs0.15_N_Samples200}.pdf}}
		\subfloat[high salinity, $c_s = 0.209$ mol/l]
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_Na_log_density_cs0.15_N_Samples200}.pdf}}
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_Cl_log_density_cs0.15_N_Samples200}.pdf}}
	\end{figure}
\end{frame}

\begin{frame}{Divalent salt. }
	\begin{columns}[T,onlytextwidth]% COLUMNS
		\column{0.55\textwidth}
			\begin{figure}
				\includegraphics[width=1\columnwidth]{figures/vmdscene.png}
			\end{figure} 		
		\column{0.4\textwidth}
			{\Large Seawater model: }
			\begin{eqnarray*}
				c_\cl&=&0.54\text{ mol/l} \\
				&&\text{of negative ions} 
				%c_\so&=&0.028 \text{ mol/l of negative divalent ions} 
			\end{eqnarray*}
			\begin{eqnarray*}
				c_\na&=&0.47\text{ mol/l} \\
				&&\text{of positive ions} 
			\end{eqnarray*}

			\begin{eqnarray*}
				c_\ca&=&0.063\text{ mol/l} \\
				&&\text{of positive divalent ions}  \\
			\end{eqnarray*}

	\end{columns}
	\begin{columns}[T,onlytextwidth]% COLUMNS
		\column{0.5\textwidth}
			\begin{eqnarray*}
				c_{\na}&\simeq0.87\cdot c_{\cl} \\
				c_{\ca}&\simeq0.117\cdot c_{\cl} \\
				%c_{\so}&\simeq&c_{\cl}\cdot0.0282/0.55=0.051c_{\cl} 
			\end{eqnarray*}
		\column{0.5\textwidth}
			\begin{eqnarray*}
				\mu_{\na}&=&\mu_{\cl} - 0.139\kT\\
				\mu_{\ca}&=&\mu_{\cl} - 2.03\kT\\
			\end{eqnarray*}	
	\end{columns}		
\end{frame}
\begin{frame}{Divalent salt. Compression.}

	\begin{figure}[t]
		\subfloat[low salinity, $c_s = 0.007$ mol/l]
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_pressure_log_density_pCa3.16_cs0.006_N_Samples200}.pdf}}
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_alpha_log_density_pCa3.16_cs0.006_N_Samples200}.pdf}}			
		\subfloat[high salinity, $c_s = 0.209$ mol/l]
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_pressure_log_density_pCa1.77_cs0.15_N_Samples200}.pdf}}
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_alpha_log_density_pCa1.77_cs0.15_N_Samples200}.pdf}}
	\end{figure}
\end{frame}

\begin{frame}{Divalent salt. Desalination or ion exchange.}
Low salinity $c_\cl = 0.007$
	\begin{figure}
	\subfloat{\includegraphics[width=0.5\columnwidth]{{figures/gel_Cl_log_density_pCa3.16_cs0.006_N_Samples200}.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{{figures/gel_Na_log_density_pCa3.16_cs0.006_N_Samples200}.pdf}}
	\vspace{0.3cm}
	\\
	\subfloat{\includegraphics[width=0.5\columnwidth]{{figures/gel_Ca_log_density_pCa3.16_cs0.006_N_Samples200}.pdf}}
	\end{figure}

\begin{frame}{Divalent salt. Desalination or ion exchange.}
High salinity $c_\cl = 0.263$
	\begin{figure}
	\subfloat{\includegraphics[width=0.5\columnwidth]{{figures/gel_Cl_log_density_pCa1.77_cs0.15_N_Samples200}.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{{figures/gel_Na_log_density_pCa1.77_cs0.15_N_Samples200}.pdf}}
	\vspace{0.3cm}	
	\\
	\subfloat{\includegraphics[width=0.5\columnwidth]{{figures/gel_Ca_log_density_pCa1.77_cs0.15_N_Samples200}.pdf}}
	\end{figure}

@@

@@

\begin{frame}{Monovalent salt. ``No electrostatics'' vs ``Mean field theory''.}
	\begin{figure}[t]
		\subfloat[low salinity, $c_s = 0.007$ mol/l]
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_pressure_log_density_lB0.0_cs0.006_N_Samples200}.pdf}}
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_alpha_log_density_lB0.0_cs0.006_N_Samples200}.pdf}}
		\subfloat[high salinity, $c_s = 0.209$ mol/l]
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_pressure_log_density_lB0.0_cs0.15_N_Samples200}.pdf}}
			{\includegraphics[width=0.5\columnwidth]{{figures/gel_alpha_log_density_lB0.0_cs0.15_N_Samples200}.pdf}}
	\end{figure}
\end{frame}
\end{document}
\begin{frame}[fragile]{Typography}
      \begin{verbatim}The theme provides sensible defaults to
\emph{emphasize} text, \alert{accent} parts
or show \textbf{bold} results.\end{verbatim}

>>>>

\section[Intro]{Introduction}
\begin{frame}[fragile]{Introduction. Desalination.}

			\begin{enumerate}
			\item Distillation
			\item Reverse osmosis
			\item Forward osmosis
			\end{enumerate}

		\begin{figure}[t]
			{\includegraphics[width=0.6\columnwidth]{{figures/RO}.pdf}}
		\end{figure}
\end{frame}
\begin{frame}[fragile]{Introduction. Forward osmsis.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]{Introduction. Manfred Wilhelm experiment}
	\begin{columns}[T,onlytextwidth]% COLUMNS
		\column{0.5\textwidth}
			\begin{figure}[t]
				{\includegraphics[width=\columnwidth]{{figures/experiment}.pdf}}
			\end{figure}

		\column{0.5\textwidth}
			\begin{figure}[t]
				{\includegraphics[width=0.9\columnwidth]{{figures/scheme}.png}}
			\end{figure}
	\end{columns}% COLUMNS

	\vspace{1cm}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]{Carnot cycle for desalination}

	\begin{columns}[T,onlytextwidth]% COLUMNS
		\column{0.3	\textwidth}
			\begin{figure}[t]
				{\includegraphics[width=0.9\columnwidth]{{figures/carnot}.pdf}}
			\end{figure}

		\column{0.7\textwidth}
		\begin{enumerate}
			\item A fully reversible desalination cycle 
			\item Analogy with Carnot cycle
		\end{enumerate}
		{\centering
			\begin{figure}[t]
				{\includegraphics[width=0.6\columnwidth]{{figures/cycles}.pdf}}
			\end{figure}
		}
		\begin{flushright}
		 Rud, et al. (2018). 
		 \end{flushright}
	\end{columns}% COLUMNS

@@

\begin{frame}{Simulation protocol.}
\begin{enumerate}
\item {\tt Choose randomly: LMD, or EX.}	\vspace{0.5cm}
\item {\tt Simulate the chosen, collecting 50 samples of: }
	\vspace{0.3cm}
	\begin{columns}[T,onlytextwidth]% COLUMNS
		\column{0.1\textwidth}
		\column{0.3\textwidth}
			\fbox{\parbox{\textwidth}{LMD: pressure, $P$, and $\{R_e\}$}}
   		\column{0.3\textwidth}
      		\fbox{\parbox{\textwidth}{EX: number of salt ions, $N_\na$ and $N_\cl$}}
		\column{0.1\textwidth}
	\end{columns} % COLUMNS
	\vspace{0.5cm}
    {\tt Check the autocorrelation of each samples array.\\
    Pearson coefficient must be < 0.2. }
	\vspace{0.5cm}
\item {\tt Repeat collecting at least 200 averages from each process}.
\end{enumerate}
\end{frame}

@@

\begin{frame}{Desalination in \{$P$, $V$\} coordinates.}
	\begin{figure}[t]
		{\includegraphics[width=\columnwidth]{{figures/results_1}.pdf}}
	\end{figure}
\end{frame}

\begin{frame}{Desalination in \{$\cs$, $V$\} coordinates.}
	\begin{figure}[t]
		{\includegraphics[width=\columnwidth]{{figures/results_3}.pdf}}
	\end{figure}
\end{frame}

\begin{frame}{Desalination in \{$N$, $V$\} coordinates.}
	\begin{figure}[t]
		{\includegraphics[width=\columnwidth]{{figures/results_4}.pdf}}
	\end{figure}
\end{frame}

\begin{frame}{Desalination in \{$\cs$, $N$\} coordinates.}
	\begin{figure}[t]
		{\includegraphics[width=\columnwidth]{{figures/results_2}.pdf}}
	\end{figure}
\end{frame}

@@

{\setbeamercolor{palette primary}{fg=black, bg=yellow}
\begin{frame}[standout]
  Questions are welcome.

  \vspace{2cm}
{\small Many thanks to }
	\begin{columns}[T,onlytextwidth]% COLUMNS
		\column{0.33\textwidth}
			\begin{figure}[t]
				{\includegraphics[width=\columnwidth]{{figures/gacr}.pdf}}\\
				\vspace{2mm}
				{\includegraphics[width=\columnwidth]{{figures/megagrant}.pdf}}\\
				\vspace{2mm}
				{\includegraphics[width=\columnwidth]{{figures/metacentrum}.png}}
			\end{figure}
			\column{0.66\textwidth}
			\centering
			\vspace{5mm}
			{\small
			\begin{itemize}
			\item Czech Science Foundation (grant 19-17847Y)			\vspace{3mm}
			\item Russian Federation, grant number 14.W03.31.0022   \vspace{3mm}
			\item “e-Infrastruktura CZ” (e-INFRA LM2018140)
			\end{itemize}}
	\end{columns}
\end{frame}
}

@@

\end{document}
\begin{frame}[fragile]{Typography}
      \begin{verbatim}The theme provides sensible defaults to
\emph{emphasize} text, \alert{accent} parts
or show \textbf{bold} results.\end{verbatim}
